---
title: "Assignment1 Joins"
author: "Chintan Nayak"
date: "08/02/2022"
output: html_document
---

```{r}
install.packages('nycflights13')
```
```{r}
library(tidyverse)
library(dplyr)
library(nycflights13)
library(ggplot2)
```
##13.4.6

##1. Compute the average delay by destination, then join on the airports data frame so you can show the spatial distribution of delays. Here’s an easy way to draw a map of the United States:

```{r}
airports %>%
  semi_join(flights, c("faa" = "dest")) %>%
  ggplot(aes(lon, lat)) +
    borders("state") +
    geom_point() +
    coord_quickmap()
```

```{r}
AvgDestDelays = flights %>%
  group_by(dest) %>%
  summarise(delay = mean(arr_delay, na.rm = TRUE)) %>%
  inner_join(airports, by = c(dest = "faa"))
```


```{r}
AvgDestDelays
```
```{r}
  ggplot(AvgDestDelays, aes(lon, lat, colour = delay)) +
  borders("state") +
  geom_point() +
  coord_quickmap()
```

##2. Add the location of the origin and destination (i.e. the lat and lon) to flights.

```{r}
airportLoc <- airports %>%
  select(faa, lat, lon)

flights %>%
  select(year:day, hour, origin, dest) %>%
  left_join(
    airportLoc,
    by = c("origin" = "faa")
  ) %>%
  left_join(
    airportLoc,
    by = c("dest" = "faa")
  )

```

##3. Is there a relationship between the age of a plane and its delays?

```{r}
plane_age = inner_join(flights,
  select(planes, tailnum, plane_year = year),
  by = "tailnum"
) %>%
  mutate(age = year - plane_year) %>%
  filter(!is.na(age)) %>%
  group_by(age) %>%
  summarise(
    dep_delay_mean = mean(dep_delay, na.rm = TRUE),
    dep_delay_sd = sd(dep_delay, na.rm = TRUE),
    arr_delay_mean = mean(arr_delay, na.rm = TRUE),
    arr_delay_sd = sd(arr_delay, na.rm = TRUE),
    n_arr_delay = sum(!is.na(arr_delay)),
    n_dep_delay = sum(!is.na(dep_delay))
  )
```


```{r}
plane_age
```
# Mean departure delay
```{r}
ggplot(plane_age, aes(x = age, y = dep_delay_mean)) +
  geom_point() +
  scale_x_continuous("Age of plane (years)", breaks = seq(0, 60, by = 10)) +
  scale_y_continuous("Mean Departure Delay (minutes)")
```
#mean arrival delay
```{r}
ggplot(plane_age, aes(x = age, y = arr_delay_mean)) +
  geom_point() +
  scale_x_continuous("Age of Plane (years)", breaks = seq(0, 60, by = 10)) +
  scale_y_continuous("Mean Arrival Delay (minutes)")
```
#There is a increase in both the delays until 10 years, then  there is a decrease in both the dalay


##4. What weather conditions make it more likely to see a delay?

```{r}
flightWeather= flights %>%
  inner_join(weather, by = c(
    "origin" = "origin",
    "year" = "year",
    "month" = "month",
    "day" = "day",
    "hour" = "hour"
  ))
```

```{r}
flightWeather
```
```{r}
flightWeather %>%
  group_by(precip) %>%
  summarise(delay = mean(dep_delay, na.rm = TRUE)) %>%
  ggplot(aes(x = precip, y = delay)) +
  geom_line() + geom_point()

```
#A delay is connected with almost any amount of precipitation.

```{r}
flightWeather %>%
  mutate(visib_cat = cut_interval(visib, n = 10)) %>%
  group_by(visib_cat) %>%
  summarise(dep_delay = mean(dep_delay, na.rm = TRUE)) %>%
  ggplot(aes(x = visib_cat, y = dep_delay)) +
  geom_point()
```
#Visibility appears to have a stronger link with delay. Delays are higher when visibility is less than 2 miles.

##5. What happened on June 13 2013? Display the spatial pattern of delays, and then use Google to cross-reference with the weather.

```{r}
flights %>%
  filter(year == 2013, month == 6, day == 13) %>%
  group_by(dest) %>%
  summarise(delay = mean(arr_delay, na.rm = TRUE)) %>%
  inner_join(airports, by = c("dest" = "faa")) %>%
  ggplot(aes(y = lat, x = lon, colour = delay)) +
  borders("state") +
  geom_point() +
  coord_quickmap()

```

